### 2018-08-02 conclusion.tex
Backup and spell check

The pipelines presented in chapter \ref{chap:pipeline} offer an output-rich and easy to deploy cost-free pieces of software. All of them have pros and cons, and their combined usage was found to provide the best results, underlining the importance of the development of programs implementing open data formats exchangeable across tools. More functionalities can be added, including \textit{de novo} and \ac{PTM} search, and missing data management. BayesQuant, the probabilistic framework for relative quantification, presented in chapter \ref{chap:model} can be passed any peptide level dataset and provide estimates of the uncertainty behind frequentist \ac{log2FC} estimates and nuisance effects. It can be customised to expand its currently implemented model. The lack of success in the attempt to model the peptide bias using the aminoacidic sequence reveals the complexity of its origin. Recently developed tools like RawQuant could be applied in future work to extract a feature-rich representation of the collected spectra that could be used to model this phenomenon with a Deep Learning approach. The results from chapter \ref{chap:benchmark} manifest that \ac{NZ} can at the same time decrease computational costs and access the state of the art in proteomics research by making use of the latest open-source software, and it is encouraged to continue on that line.

>>>>

The pipelines presented in chapter \ref{chap:pipeline} offer an output-rich and easy to deploy cost-free pieces of software. All of them have pros and cons, and their combined usage was found to provide the best results, underlining the importance of the development of programs implementing open data formats exchangeable across tools. More functionalities can be added, including \textit{de novo} and \ac{PTM} search, fractioned samples and missing data management. BayesQuant, the probabilistic framework for relative quantification, presented in chapter \ref{chap:model}, can be passed any peptide level dataset and provide estimates of the uncertainty beyond frequentist \ac{log2FC} estimates. It can be customized to expand its currently implemented model. The lack of success in the attempt to model the peptide bias using the aminoacidic sequence reveals the complexity of its origin. Recently developed tools  could be applied in future work to extract a feature-rich representation of the collected spectra that could be used to model this phenomenon with a Deep Learning approach. The results from chapter \ref{chap:benchmark} manifest that \ac{NZ} can at the same time decrease computational costs and access the state of the art in proteomics research by making use of the latest open-source software, and it is encouraged to continue on that line.

### 2018-08-02 modelling.tex
Backup and spell check

The Bayesian statistical framework provides an alternative point of view by revolving the role played by the model parameters and the observed data. While frequentist statistics treats the data as random and the parameters as fixed, a Bayesian framework swaps their roles and yields a probability distribution for any model parameter given the provided data. Bayesian analyses rely on the observed data and prior knowledge about the phenomenon only, and comes equipped with an equation that mathematically formalizes how to introduce these two dependencies in a valid way. This is the so called Bayes\textquotesingle s theorem.

@@

The data is thus said to be affected by different effects acting independently, and in order to model them properly, multilevels need to be defined. It can be done via three approaches: pooled, unpooled, and hierarchically (see figure \ref{fig:multilevel}). Hierarchical modelling is used when the sampling variance is not the only source of variation among the parameters, but they still share a dependency. The intercept and the three effects were modelled as independent hierarchies.

@@

However the posterior distribution is obtained, a proper Bayesian analysis is not finished without a model checking step that looks at potential problems happenning in the fitting step, and confirms the goodness of fit of the model to the data.

@@

PyMC3 is a mature Python module dedicated to support probabilistic programming. It features   intuitive model specification syntax, modern and powerful \ac{MCMC} sampling algorithms like the No-U-Turn-Sampler (\ac{NUTS}) as well as Automatic Differentiation Variational Inference (ADVI) \footnote{\href{https://github.com/pymc-devs/pymc3}{https://github.com/pymc-devs/pymc3}}. Therefore, it provides the tools required to build a Bayesian model for the probabilisitc estimation of protein \ac{log2FC} in proteomics datasets, that is, together with an estimate of its uncertainty.

@@

The hierarchical structure of the model is set by the definition of a parameter distribution from which the value for each element (peptide) being modelled is sampled. This can be done by defining a new normal distribution where its $\mu$ and $\sigma$ are set to the random variables defined above. Im PyMC3 code,

@@

\begin{minted}[mathescape,
               linenos,
               numbersep=5pt,
               obeytabs=true,tabsize=2,
               frame=lines,
               firstnumber=last,
               framesep=2mm]{python}
# VI approximation: much faster and pretty accurate
trace_advi = bayesquant.fit(model_name=p, n_draws=40000)
\end{minted}

@@

%\begin{figure}[H]
% \centering 
% \includegraphics{hierarch_diagram.pdf}
% \caption{}
% label{fig:dbda}
%\end{figure}
\begin{figure}[!h]
\input{hierarch_diagram}
\caption{Kruschke-like diagram of the Bayesian model}
\label{fig:dbda}
\end{figure}

@@

The 10k traces produced by BayesQuant from the \ac{VI} approximation function exhibited a clear random walk behaviour, which decreases the posibility that the approximation was wrong. Moreover, the  effects caused by the different peptides and runs were estimated and separated from the overall observed noise, allowing for a better estimation of the \ac{log2FC} (see figure \ref{fig:traceplots}).

@@

The study of the \ac{log2FC} posterior probability distribution (see figure \ref{fig:posteriors}) shows the most likely parameter values and empowers Bayesian Null Hypothesis Testing. The 95\% High Probability Density Interval (\ac{HPDI}), which is the narrowest interval containing 95\% of the total probability, indicates the most likely true values of the parameter. Presence or absence of overlap between the 95\% \ac{HPDI} and the Region Of Practical Equivalence (\ac{ROPE}) provides a formal way of discarding a point value of the parameter. The \ac{ROPE} is a small interval considered to be essentually the same as the null value \cite{Kruschke}. In this and all future applications in the thesis, the null value is set to 0, and the \ac{ROPE} is arbitrality set to $[-0.4, 0.4]$ based on the data.

@@

The posterior predictive checks on both proteins pinpoints that the fitted models captured a data generation process approximating what was observed (see figure \ref{fig:ppc}). The three observed peptides were found to be undistinguishable from the 500 simulated ones.

@@

The consistent observation that \textit{Homo sapiens} proteins got a \ac{log2FC} estimate around 0, and \textit{E. coli} proteins got an estimate which overlapped the predefinned \ac{ROPE} in only three cases (two of which on proteins with two peptides), confirmed the validity of the quantification framework.

@@

Given the open-source nature of the program, it is possible to build upon it and extend its functionality. The possibility of making use of the peptide sequences to model the peptide effect using a linear model was considered and put to practice. The code was thus extended with the following snippets of code:

@@

\begin{equation}
\nonumber \sigma^{\theta} \sim \mathcal{HN}(1)
\end{equation}
\begin{equation}
\nonumber \theta_k \sim \mathcal{N}(0, \sigma^{\theta})
\end{equation}
\begin{equation}
\nonumber \mu^p \sim \mathcal{N}(\sum_{k=0}^{k=K}{x_k \theta_k}, \sigma^p)
\end{equation}

@@

@@

A solid way of checking the results are not flawed is the specification of alternative priors. Unless precise prior knowledge is available, the selection of one prior over another prior to model the beliefs about the system should not have any impact on the results. It is for that reason that models built upon different priors could be run to further validate the program.

@@

Besides the predefinition of a \ac{ROPE} and the check of overlap with the 95 \% \ac{HPDI}, more advanced model comparison methods making use of the Bayes factor are available. The Bayes factor is defined as the ratio of the posterior probability of the models given the data \cite{Kruschke}. It could be used to measure how much more likely a model taking into account a treatment effect is compared to another one where the treatment effect is considered null i.e the \ac{log2FC} is set to 0. The result would provide an alternative criterion for the declaration of a protein being differentially abundant.

@@

\item The characterization of powerful sequence properties that specifically capture differential ionization efficiency is undone work, as most feature extractors are designed to solve different problems, like protein sructure prediction.

@@

Modelling such a complex phenomenon will require accordingly complex predictive architectures and the introduction of proper neural networks, capable of deciphering the intricate pattern mapping sequence to noise in the spectrometer\textquotesingle s measurements.

@@

A Bayesian framework for the relative quantification of protein abundance rations (Fold-Change, FC) using \ac{MS1} intensity was presented in this chapter. Unlike the currently prevailing methods in the field, it gives uncertainty estimates that provide a direct interpretation about the accuracy of the quantification. Moreover, preliminary results on the sequence modelling of the peptide bias revealed that more complex architectures and sequnce features are required to model noise with peptide sequences. Finally, the method can be used to investigate the presence of batch effects in \ac{MS} experiments for their minimisation in future experiments. While the software can be integrated in any pipeline producing peptide-level data, further testing and improvements are probably required for optimal performance.

>>>>

The Bayesian statistical framework provides an alternative point of view by revolving the role played by the model parameters and the observed data. While frequentist statistics treats the data as random and the parameters as fixed, a Bayesian framework swaps their roles and yields a probability distribution for any model parameter given the provided data. Bayesian analyses rely on the observed data and prior knowledge about the phenomenon only, and come equipped with an equation that mathematically formalizes how to introduce these two dependencies in a valid way. This is the so called Bayes\textquotesingle s theorem.

@@

The data is thus said to be affected by different effects acting independently, and in order to model them properly, multi-levels need to be defined. It can be done via three approaches: pooled, unpooled, and hierarchically (see figure \ref{fig:multilevel}). Hierarchical modelling is used when the sampling variance is not the only source of variation among the parameters, but they still share a dependency. The intercept and the three effects were modelled as independent hierarchies.

@@

However the posterior distribution is obtained, a proper Bayesian analysis is not finished without a model checking step that looks at potential problems happening in the fitting step, and confirms the goodness of fit of the model to the data.

@@

PyMC3 is a mature Python module dedicated to support probabilistic programming. It features   intuitive model specification syntax, modern and powerful \ac{MCMC} sampling algorithms like the No-U-Turn-Sampler (\ac{NUTS}) as well as Automatic Differentiation Variational Inference (ADVI) \footnote{\href{https://github.com/pymc-devs/pymc3}{https://github.com/pymc-devs/pymc3}}. Therefore, it provides the tools required to build a Bayesian model for the probabilistic estimation of protein \ac{log2FC} in proteomics datasets, that is, together with an estimate of its uncertainty.

@@

The hierarchical structure of the model is set by the definition of a parameter distribution from which the value for each element (peptide) being modelled is sampled. This can be done by defining a new normal distribution where its $\mu$ and $\sigma$ are set to the random variables defined above. In PyMC3 code,

@@

\begin{minted}[mathescape,
               linenos,
               numbersep=5pt,
               obeytabs=true,tabsize=2,
               frame=lines,
               firstnumber=last,
               framesep=2mm]{python}
# VI approximation: much faster and accurate
trace_advi = bayesquant.fit(model_name=p, n_draws=40000)
\end{minted}

@@

%\begin{figure}[H]
% \centering 
% \includegraphics{hierarch_diagram.pdf}
% \caption{}
% label{fig:dbda}
%\end{figure}
\begin{figure}[!h]
\input{hierarch_diagram}
\caption[BayesQuant Kruschke-like diagram]{\textbf{BayesQuant Kruschke-like diagram}. Extends figure \ref{fig:daft_model} to give more detail into the model. Prior probability distributions are indicated with "distrograms", and the dependencies between them via arrows. The hyperparameters are represented by latin capital letters.}
\label{fig:dbda}
\end{figure}

@@

The 10k traces produced by BayesQuant from the \ac{VI} approximation function exhibited a clear random walk behaviour, which decreases the possibility that the approximation was wrong. Moreover, the  effects caused by the different peptides and runs were estimated and separated from the overall observed noise, allowing for a better estimation of the \ac{log2FC} (see figure \ref{fig:traceplots}).

@@

The study of the \ac{log2FC} posterior probability distribution (see figure \ref{fig:posteriors}) shows the most likely parameter values and empowers Bayesian Null Hypothesis Testing. The 95\% High Probability Density Interval (\ac{HPDI}), which is the narrowest interval containing 95\% of the total probability, indicates the most likely true values of the parameter. Presence or absence of overlap between the 95\% \ac{HPDI} and the Region Of Practical Equivalence (\ac{ROPE}) provides a formal way of discarding a point value of the parameter. The \ac{ROPE} is a small interval considered to be essentially the same as the null value \cite{Kruschke}. In this and all future applications in the thesis, the null value is set to 0, and the \ac{ROPE} is arbitrarily set to $[-0.4, 0.4]$ based on the data.

@@

The posterior predictive checks on both proteins pinpoint that the fitted models captured a data generation process approximating what was observed (see figure \ref{fig:ppc}). The three observed peptides were found to be indistinguishable from the 500 simulated ones.

@@

The consistent observation that \textit{Homo sapiens} proteins got a \ac{log2FC} estimate around 0, and \textit{E. coli} proteins got an estimate which overlapped the predefined \ac{ROPE} in only three cases (two of which on proteins with two peptides), confirmed the validity of the quantification framework.

@@

The possibility of making use of the peptide sequences to model the peptide effect using a linear model was considered and put to practice. The code was thus extended with the following snippets of code:

@@

\begin{equation}
\nonumber \sigma^{\theta} \sim \mathcal{HN}(1)
\end{equation}
\begin{equation}
\nonumber \theta_k \sim \mathcal{N}(0, \sigma^{\theta})
\end{equation}
\begin{equation}
\nonumber \mu^p \sim \mathcal{N}(\sum_{k=1}^{k=K}{x_k \theta_k}, \sigma^p)
\end{equation}

@@

A list of points to address to improve how BayesQuant works is discussed in this section.

@@

A solid way of checking the results are not flawed is the specification of alternative priors. Unless precise prior knowledge is available, the selection of one prior over another to model the beliefs about the system should not have great impact on the results. It is for that reason that models built upon different priors could be run to further validate the program. Moreover, a formal posterior validation beyond visual inspection of simulated data could be implemented.

@@

Likewise, a formal way of establishing the \ac{ROPE} in the quantification context could be enacted depending on the data. On the other hand, a more advanced model comparison method could be built upon the Bayes factor. The Bayes factor is defined as the ratio of the posterior probability of two models given the data \cite{Kruschke}. It could be used to measure how much more likely a model taking into account a treatment effect is compared to another one where the treatment effect is considered null i.e the \ac{log2FC} is set to 0. The result would provide an alternative criterion for the declaration of a protein being differentially abundant.

@@

\item The characterization of powerful sequence properties that specifically capture differential ionization efficiency is undone work, as most feature extractors are designed to solve different problems, like protein structure prediction.

@@

Modelling such a complex phenomenon will require accordingly complex predictive architectures and the introduction of proper neural networks, capable of deciphering the intricate pattern mapping sequence to noise in the spectrometer\textquotesingle s measurements. The development of very new RAW spectra-files parsing-tools like RawQuant \cite{Kovalchik2018} could be used in a future implementation to provide the required feature-rich data.

@@

A Bayesian framework for the relative quantification of protein abundance ratios (Fold-Change, FC) using \ac{MS1} intensity was presented in this chapter. Unlike the currently prevailing methods in the field, it gives uncertainty estimates that provide a direct interpretation about the accuracy of the quantification. Moreover, preliminary results on the sequence modelling of the peptide bias revealed that more complex architectures and sequence features are required to model noise with peptide sequences. Finally, the method can be used to investigate the presence of batch effects in \ac{MS} experiments for their minimisation in future experiments. While the software can be integrated in any pipeline producing peptide-level data, further testing and improvements are probably required for optimal performance.

### 2018-08-02 msreview.tex
Backup and spell check

The main source of data in proteomics is \ac{MS}. However, different approaches to how this technique is used make for two paradigms in proteomics analysis: top-down and bottom-up \cite{Joshi2016}. In the top-down paradigm, intact proteins are directly used for the analysis \cite{Sinitcyn2018}. In the bottom-up paradigm (see figure \ref{fig:proteomics_overview}), the proteins are first cleaved into smaller parts, and these parts are then used for identification, characterization, and quantification. These smaller parts are peptides \cite{Barsnes2008}, consisting of around 10 chained aminoacids. Such peptides acquire physicochemical properties fitting the requirements of the downstream analytical methods, mainly the mass spectrometer (MS), which performs the data acquisition. The bottom-up paradigm is most often used because peptides are much more suitable to analysis by mass spectrometry, as illustrated in section \ref{subsec:the_detector}. This review will focus on the Data-Dependent Aquisition (\ac{DDA} approach within bottom-up proteomics, which currently is the most frequent workflow in proteomics \cite{Sinitcyn2018}. In this approach, the equipment is configured for the analysis of one peptide at a time, with the drawback of being biased for the most abundant peptides.

@@

An \ac{MS} experiment starts with the generation of a protein sample from the biological system of interest. Proteins are then denaturalized  so as to remove bias due to the divergent properties acquired by folded proteins. Then, proteins are subjected to digestion with specific proteases i.e protein-cutting molecules, which cut the aminoacidic chains following a predictable pattern. Trypsin is the most frequently used protease. It cuts peptidic bonds whenever a positively charged residue, either Lysine (K) or Arginine (R), lies on the carboxyl side of the peptidic bond. Since roughly 1/10 residues are either R or K, the average peptide length is 10 residues, as mentioned before. As demonstrated in section \ref{subsec:the_detector}, this length distribution is fitted to the resolution of the MS analyzer. Moreover, in as much R and K are positively charged aminoacids (see figure \ref{fig:aminoacids}), the resulting peptide is guaranteed in most cases to be able to capture at least one charge, which is key in the \ac{MS} workflow as described in section \ref{sec:the_mass_spectrometer}. All of these properties combined, together with its low prize, makes Trypsin the protease of choice in this step for most cases.

@@

Missed cleavages can happen due to steric impediments or the presence of specific aminoacids that can weaken the enzyme\textquotesingle s function \cite{Siepen2007}. This is the case of Trypsin whenever the residue on the other side of the peptidic bond is Proline. Altogether, a variability is created in the cleavage process that, though limited, needs to be taken care of in downstream analysis, as it could introduce biases in peptide observability.

@@

If presented with the problem of analyzing a mixture of peptides, the capacities of mass spectrometers are easily overwhelmed by a too complex mixture, resulting in the analysis of only a minor part of the total protein of the sample \cite{Barsnes2008}. This can be surmounted by analyzing one peptide in the sample at a time. The required sample separation is achieved by High-Performance Liquid Cromatography (\ac{HPLC}) methods, like reverse phase chromatography (separating on hydrophobicity) and strong cation exchange chromatography (separating on isoelectric point) \cite{Barsnes2008}.

@@

All mass spectrometers exploit the physical properties of mass and electric charge exhibited by the analyzed components. Ionization of the analytes is absolutely essential prior to any measurement, as analytes left uncharged will be unobservable to the equipment.

@@

The acquired charge yields a mass/charge (\ac{m/z}) ratio, a property that is applied in the downstream component separation and measurement steps. It is measured in $\frac{Da}{e}$, sometimes referred to as Th (Thomson).

@@

Moreover, two other analyzers exist which combine mass analysis with intensity measurement. These are Fourier Transform Ion Cyclotron Resonance (\ac{FT-ICR}) and Orbitrap. They both register cylotron resonance frequencies that are Fourier transformed into the spectrum space. Remarkably, \ac{FT-ICR} exhibits great resolving power, at the cost of high maintenance costs and difficult operability \cite{Barsnes2008}.

@@

On the other hand, due to the natural occurrence of isotopes, particularly \ce{^{13}_{}C}, the same peptide will induce the measurement of several signals with very close \ac{m/z} values. They constitute the isotopic envelope of the ion (see figure \ref{fig:envelope}) \cite{Mirzaei2016}, and represent the signal created by peptides containing an increasing number of \ce{^{13}_{}C} atoms or other naturally occurring isotopes. Every time a \ce{^{12}_{}C} is replaced by \ce{^{13}_{}C}, the mass increases by 1 Da. Even though the natural abundance of \ce{^{13}_{}C} is 1.1 \%, the sheer number of carbon atoms in a peptide makes it likely that at least one or even more carbon atoms will be \ce{^{13}_{}C}, eventually driving the pure \ce{^{12}_{}C} signal to comparatively small intensity values, and down to intensities below the background noise. Such event can be problematic if it entails that the \ce{^{13}_{}C} peak is confused for the \ce{^{12}_{}C} peak.

@@

Computational analysis of MS data starts with the matching of the spectra to a reference proteome (see figure \ref{fig:computational_analysis}). \ac{MS} search engines are capable of performing the crucial step of peptide to spectrum matching (PSM). During this step, search engines attempt to match each observed spectrum to theoretical spectra based on the reference proteome. Theoretical spectra are obtained via (I) \textit{in silico} simulation of the spectra produced by the peptides contained in the proteome, in turn obtained by performing a virtual cleaveage of the sequences in the proteome, or (II) spectral libraries, consisting of annotated observed spectra \cite{Li2012}.

@@

\item \textbf{Peptide inference}: infer the peptides present in the sample.
\item \textbf{Protein inferece \textit{proper}}: based on the inferred peptides, infer what proteins generated them. This is not trivial as peptides are degenerate and frequently map to more than one protein \cite{Li2012}.
\end{enumerate}

@@

Quantification can be spectrum counting (\ac{SC}) or extractde ion currents (\ac{XIC}) based.

>>>>

The main source of data in proteomics is \ac{MS}. However, different approaches to how this technique is used make for two paradigms in proteomics analysis: top-down and bottom-up \cite{Joshi2016}. In the top-down paradigm, intact proteins are directly used for the analysis \cite{Sinitcyn2018}. In the bottom-up paradigm (see figure \ref{fig:proteomics_overview}), the proteins are first cleaved into smaller parts, and these parts are then used for identification, characterization, and quantification. These smaller parts are peptides \cite{Barsnes2008}, consisting of around 10 chained aminoacids. Such peptides acquire physicochemical properties fitting the requirements of the downstream analytical methods, mainly the mass spectrometer (MS), which performs the data acquisition. The bottom-up paradigm is most often used because peptides are much more suitable to analysis by mass spectrometry, as illustrated in section \ref{subsec:the_detector}. This review will focus on the Data-Dependent Acquisition (\ac{DDA} approach within bottom-up proteomics, which currently is the most frequent workflow in proteomics \cite{Sinitcyn2018}. In this approach, the equipment is configured for the analysis of one peptide at a time, with the drawback of being biased for the most abundant peptides.

@@

An \ac{MS} experiment starts with the generation of a protein sample from the biological system of interest. Proteins are then denatured  so as to remove bias due to the divergent properties acquired by folded proteins. Then, proteins are subjected to digestion with specific proteases i.e protein-cutting molecules, which cut the aminoacidic chains following a predictable pattern. Trypsin is the most frequently used protease. It cuts peptidic bonds whenever a positively charged residue, either lysine (K) or arginine (R), lies on the carboxyl side of the peptidic bond. Since roughly 1/10 residues are either R or K, the average peptide length is 10 residues, as mentioned before. As demonstrated in section \ref{subsec:the_detector}, this length distribution is fitted to the resolution of the MS analyzer. Moreover, in as much R and K are positively charged aminoacids (see figure \ref{fig:aminoacids}), the resulting peptide is guaranteed in most cases to be able to capture at least one charge, which is key in the \ac{MS} workflow as described in section \ref{sec:the_mass_spectrometer}. All of these properties combined, together with its low prize, makes Trypsin the protease of choice in this step for most cases.

@@

Missed cleavages can happen due to steric impediments or the presence of specific aminoacids that can weaken the enzyme\textquotesingle s function \cite{Siepen2007}. This is the case of Trypsin whenever the residue on the other side of the peptidic bond is proline. Altogether, a variability is created in the cleavage process that, though limited, needs to be taken care of in downstream analysis, as it could introduce biases in peptide observability.

@@

If presented with the problem of analyzing a mixture of peptides, the capacities of mass spectrometers are easily overwhelmed by a too complex mixture, resulting in the analysis of only a minor part of the total protein of the sample \cite{Barsnes2008}. This can be surmounted by analyzing one peptide in the sample at a time. The required sample separation is achieved by High-Performance Liquid Chromatography (\ac{HPLC}) methods, like reverse phase chromatography (separating on hydrophobicity) and strong cation exchange chromatography (separating on isoelectric point) \cite{Barsnes2008}.

@@

All mass spectrometers exploit the physical properties of mass and electric charge exhibited by the analyzed components. Ionization of the analytes is absolutely essential prior to any measurement, as analytes left uncharged will be invisible to the equipment.

@@

The acquired charge yields a mass/charge (\ac{m/z}) ratio, a property that is applied in the downstream component separation and measurement steps. It is measured in $\frac{Da}{e}$ (Dalton over elementary charge), sometimes referred to as Th (Thomson).

@@

Moreover, two other analyzers exist which combine mass analysis with intensity measurement. These are Fourier Transform Ion Cyclotron Resonance (\ac{FT-ICR}) and Orbitrap. They both register cyclotron resonance frequencies that are Fourier transformed into the spectrum space. Remarkably, \ac{FT-ICR} exhibits great resolving power, at the cost of high maintenance costs and difficult operation \cite{Barsnes2008}.

@@

On the other hand, due to the natural occurrence of isotopes, particularly \ce{^{13}_{}C}, the same peptide will induce the measurement of several signals with very close \ac{m/z} values. They constitute the isotopic envelope of the ion (see figure \ref{fig:envelope}) \cite{Mirzaei2016}, and represent the signal created by peptides containing an increasing number of \ce{^{13}_{}C} atoms or other naturally occurring isotopes. Every time a \ce{^{12}_{}C} is replaced by \ce{^{13}_{}C}, the molecule\textquotesingle s mass increases by 1 Da. Even though the natural abundance of \ce{^{13}_{}C} is 1.1 \%, the sheer number of carbon atoms in a peptide makes it likely that at least one or even more carbon atoms will be \ce{^{13}_{}C}, eventually driving the pure \ce{^{12}_{}C} signal to comparatively small intensity values, and down to intensities below the background noise. Such event can be problematic if it entails that the \ce{^{13}_{}C} peak is confused for the \ce{^{12}_{}C} peak.

@@

Computational analysis of MS data starts with the matching of the spectra to a reference proteome (see figure \ref{fig:computational_analysis}). \ac{MS} search engines are capable of performing the crucial step of peptide to spectrum matching (PSM). During this step, search engines attempt to match each observed spectrum to theoretical spectra based on the reference proteome. Theoretical spectra are obtained via (I) \textit{in silico} simulation of the spectra produced by the peptides contained in the proteome, in turn obtained by performing a virtual cleavage of the sequences in the proteome, or (II) spectral libraries, consisting of annotated observed spectra \cite{Li2012}.

@@

\item \textbf{Peptide inference}: infer the peptides present in the sample.
\item \textbf{Protein inference \textit{proper}}: based on the inferred peptides, infer what proteins generated them. This is not trivial as peptides are degenerate and frequently map to more than one protein \cite{Li2012}.
\end{enumerate}

@@

Quantification can be spectrum counting (\ac{SC}) or extracted ion currents (\ac{XIC}) based.
